### 2019-04-07 papers/journal/encoding.tex
Font problems fixed
Section 6 completed

%\externaldocument{sat-table}
%\externaldocument{smt-table}
%\externaldocument{qbf-table}
\noindent In the previous section, we have discussed four problems, $\fbmc$ search, $\fsmt$ search, $\fqbf$ search and synthesis problem.
%
In this section we will present encodings for each of them.
%
Although the underlying technique for $\fbmc$ and $\fsmt$ search varies, the encoding is very similar. 
%
So, instead of presenting serially the complete encoding of both, 
%$\fsmt$ search, $\fbmc$ search, $\fqbf$ search and synthesis problems.
%
we will start with the complete $\fsmt$ search encoding and present only the relevant part of the the $\fbmc$ encoding to show a contrast.
%contrast with the $\fbmc$ encoding.

@@

%\subsection{Boolean satisfiability of the search problem}
\noindent We translate the search problem into a Boolean satisfiability
problem and use SMT solvers to find the satisfying VTSs.
%
We will first present the variables used to encode the
VTSs and the properties.
%
Afterwards, we will present the constraints corresponding to the
properties.

\subsubsection{Variables for VTS description}
%
We assume that size of the graph is $\nu$ and number of molecules is
$\mu$.
%
Furthermore, we also limit the maximum number $\pi$ of edges present
between two nodes.
%
Here, we list the vector of Boolean variables and uninterpreted function symbols
that encode the VTSs.
\begin{enumerate}

@@

\subsubsection{Constraints on presence, activity of the molecule, and pairing matrix}
%
We need the following constraints \texttt{EdgeC},
\texttt{ActivityC} and \texttt{PairingC} to encode the basic structure of VTSs.
%
For an edge to exist it should have one molecule present. 
%
\begin{align*}
  \texttt{PresentE} = \bigwedge\limits_{i,j,q} (\bigvee_m e_{i,j,q,m} )\limplies e_{i,j,q}
\end{align*}
The edge labels are subset of the node labels of the source and target nodes.
\begin{align*}
  \texttt{LabelE} = \bigwedge\limits_{i,j,q,m} e_{i,j,q,m} \limplies (n_{i,m} \land n_{j,m} )
\end{align*}
Self edges are not allowed. 
\begin{align*}
  \texttt{SelfE} = \bigwedge\limits_{i,q} \neg e_{i,i,q}
\end{align*}
\noindent We define \texttt{EdgeC} to collect all the edge related constraints. 
%
\texttt{EdgeC} states that each edge has at least one molecule,
 edge labels are consistent with node labels and there are no self loops.    
\begin{align*}
  \texttt{EdgeC} = \texttt{PresentE} \land \texttt{LabelE}  \land \texttt{SelfE} 
\end{align*}
\noindent If a molecule is active on an edge, it should be present on the edge.
%
\begin{align*}
  \texttt{ActiveMolE} = \bigwedge\limits_{i,j,q,m} b_{i,j,q,m} \limplies e_{i,j,q,m}\label{eq:f1}
\end{align*}
A molecule should be present to be active on a node.  
\begin{align*}
    \texttt{ActiveMolN} = \bigwedge\limits_{i,m} a_{i,m} \limplies n_{i,m}
\end{align*}
\noindent We define \texttt{ActivityC} to collect all the molecule activity related constraints. 
\texttt{ActivityC} states that all active molecules are present.
\begin{align*}
  \texttt{ActivityC} = \texttt{ActiveMolE} \land \texttt{ActiveMolN} 
\end{align*}
\noindent We divide the pairing matrix into Q-SNAREs and R-SNAREs blocks and set the dependency bits of R-SNAREs to R-SNAREs in the pairing matrix to be all 0's.
%It reduces the search space.
% Condition on $p_{kk'}$. \ashu{why???} .
%\begin{align}
%  \bigwedge\limits_{(x < M/2 \, \land  \, y < M/2) \lor  (x >= M/2 \land y >= M/2)} \neg p(x,y)
%  \tag{V6}\label{eq:c3}
%\end{align}

@@

\subsubsection{Well-fused constraints}
Constraint \texttt{Fusion1} encodes that each edge must fuse with
its destination node.
%
Constraint \texttt{Fusion2} encodes that each edge should not
be able to potentially fuse with any node other than its destination node.
\begin{align*}
\texttt{Fusion1} =\;&
\bigwedge\limits_{i,j,q} e_{i,j,q} \limplies
\bigvee_{{m_1,m_2,m_3,m_4}} (\Land_{l=1}^4 ( b_{i,j,q,m_l} \lor a_{j,m_l} ) \land 
\Lor\limits_{l=1}^4 b_{i,j,q,m_l} \land\\
& \qquad \qquad \qquad \qquad \Lor\limits_{l=1}^4 a_{j,m_l}\land p_{\{m_1,m_2,m_3,m_4\}})
\\
\texttt{Fusion2} =\;&
\bigwedge\limits_{\mathclap{i,j,q,m_1,m_2,m_3, l \in \{1,..,3\}}} b_{i,j,q,m_1} \land ..\land b_{i,j,q,m_l} \limplies \\
& \hspace{2cm} \neg 
\bigvee_{\mathclap{j \neq j^{\prime}, m_{l+1}^{\prime},..,m_{4}^{\prime}}} ( a_{j^{\prime},m_{l+1}^{\prime}} \land .. \land a_{j^{\prime},m_4^{\prime}} \land p_{\{m_1,..,m_l,m^{\prime}_{l+1},..,m^{\prime}_4\}})
\\
\end{align*}
We define \texttt{FusionC} to collect all the fusion related constraints.
% 
\texttt{FusionC} states that every edge in the graph is well-fused.
\begin{align*}
\texttt{FusionC} = \texttt{Fusion1} \land \texttt{Fusion2} 
\end{align*}
% Fusion rules consist of two different mechanisms.
% \begin{enumerate}
% \item  A \textbf{pairing mechanism} which determines compatible Q-R pairs on vesicles and compartments that can cause fusion.
% \item \textbf{regulatory mechanisms} on the edges and nodes (possibly distinct) which regulates molecules activity based on the presence/absence of other molecules on the corresponding node or edge.
% \end{enumerate}
% Boolean constraint~\eqref{eq:fuse1} and~\eqref{eq:fuse2} are used to encode well-fused
% property.

@@

\subsubsection{Regulation on nodes and edges (Activity function)}
We are considering several models of VTS that differ in types of regulations on the molecule both on the node and on the edge.  
%constraints on the activity of molecules.
%
%We will present~\eqref{eq:ann}-~\eqref{eq:aep} that encodes
%the varying constraints.
%
The activity of the molecules are controlled by activity functions.
%
We will have encoded five variation of the Boolean functions based on the biological discussion.
%
The concrete functions encoding is presented below.
%
We will use \texttt{NodeFun}$_m$ and \texttt{EdgeFun}$_m$ for node and edge regulations respectively.
%

The following constraint encodes no conditions on activities on nodes,
i.e., all the present molecules on the nodes are active.
\begin{align*}
\texttt{FunOnN} =& \bigwedge\limits_{i,m} n_{i,m} = a_{i,m}  
\end{align*}
The following constraint encodes that activity of a molecule $m$ on the node is
defined by a Boolean function $\nodef_m$ of presence of molecules present on that node.
\begin{align*}
\texttt{FunBfN} =& \bigwedge\limits_{i,m} n_{i,m} \limplies a_{i,m} =  \nodef_m (n_{i,1},\dots,n_{i,\mu}) 
\end{align*}
The following constraint encodes no conditions on activities on edges,
i.e., all the present molecules on the edges are active.
\begin{align*}
 \texttt{FunOnE} =& \bigwedge\limits_{i,j,q,m} e_{i,j,q,m} = b_{i,j,q,m}
\end{align*}
The following constraint encodes that activity of a molecule $m$ on the edge is
defined by a Boolean function $\edgef_m$ of presence of molecules present on that edge.
\begin{align*}
 \texttt{FunBfE} =& \bigwedge\limits_{i,j,q,m} e_{i,j,q,m} \limplies b_{i,j,q,m} = \edgef_k(e_{i,j,q,1}, .., e_{i,j,q,\mu} )
\end{align*}
%
The following constraint encodes that the activity of the molecules on
edges is defined by inhibition by other molecules based on the pairing
matrix. 
\begin{align*}
\texttt{FunPairE} =&	\begin{aligned}[t]
\bigwedge\limits_{i,j,q, m}  [e_{i,j,q,m} \land \bigwedge_{m_l \neq m,l \in \{1,..,3\}} e_{i,j,q,m'} \land p_{m, m_l}] \limplies\\
	(\neg b_{i,j,q,m} \land \bigwedge_{\mathclap{m_l \neq m,l, p_{m, m_l} \in \{1,..,3\}}} \neg b_{i,j,q,m_l})\end{aligned}\\
\end{align*}
%
We denote the particular choice of node and edge function combined by \texttt{ActivityFun}.
\begin{align*}
\texttt{ActivityFun} = \texttt{NodeFun}_m \land \texttt{EdgeFun}_m
\end{align*}
\subsubsection{Constraints for stability condition}
%
We use $m$-reachability to encode the stability condition in VTSs.
%
The following constraint recursively encodes that node $j$ is $m$-reachable from node $i$ in less than $p$ steps
if either there is a direct edge between $i$ and $j$ with $m$ present on the edge or there is a edge between $i^{\prime \prime}$ and
$(i \neq i^{\prime \prime})$ with $m$ present, and j is $m$-reachable from $i'$ in less than $p-1$ steps.
%
\begin{align*}
  \texttt{Paths}(r) &= \bigwedge\limits_{i,j,m,p} r_{i,j,m,p} \limplies (\bigvee_{q} \, e_{i,j,q,m} \lor \bigvee_{i\neq i^{\prime}} ( \, \bigvee_{q} e_{i,i^{\prime},q,m}) \land r_{i^{\prime},j,m,p-1} )
\end{align*}
Now we can encode stability using the reachability variables.
and say if there is an $m$-edge between $i$ and $j$, there is
$m$-reachable path from $j$ to $i$.
\begin{align*}
 \texttt{Loop}(r) &= \bigwedge\limits_{i,j,m} (\bigvee_{q} e_{i,j,q,m}) \limplies r_{j,i,m,\nu}
\end{align*}
We define \texttt{StabilityC} to collect all the stability related constraints.
% 
\texttt{StabilityC} states that every molecule is moving in a cycle.
\begin{align*}
\texttt{StabilityC} = \texttt{Paths}(r) \land \texttt{Loop}(r) 
\end{align*}
% Boolean variable
% F3 and F2 are used to model these constraints.

@@

\noindent Our first attempt was to encode the existential condition as a $\fbmc$ search problem~\cite{shukla2017discovering}.
%
%In this section we will present the $\fbmc$ encoding of the same.
%
%Although the techniques used for $\fbmc$ and $\fsmt$ is very different because we are 
%In $\fbmc$ we only partially unroll the program to check for the solutions within that subset. 
%in the case of $\fbmc$, the encoding is not considerably different.
% 
The main difference between $\fsmt$ and $\fbmc$ is the encoding of 
%I n comparison to $\fsmt$, the $\fbmc$ search encodes 
reachability (both for the stability and connectivity).  
%]
Here, we will only present the stability encoding.
%the rest of the encodings are similar to the one presented in the previous section, Section~\ref{enc:smt}.

@@

% with each bit corresponding to the presence or absence of the molecule. 
%
%In the $\fbmc$ encoding, variables are represented as an array of bit vectors of size $M$. 
%
%We initialize the system with a fixed number of node size (N) and distinct molecular types (M). 
%%
%We use a two-dimensional array variable \textit{Graph}[N][N] to represent the transport graph. The graph variable is set to x, Graph[i][j] = x, if there are exactly x edges between node i and j. 
%\begin{align*}
%graph [i] [j] = x.
%\end{align*}
%
%For example, every edge is represented as a bit vector of size $M$ with each bit corresponding to the presence or absence of the molecule. 
%
%Bit vector \texttt{bvEdge}$[i][j][q]$ represent the $q$th edge between node $i$ and $j$.
%
%The index of the bit vector implicitly determines the corresponding molecule we are referring to in the edge.
% 
%The molecule set is divided into Q and R-SNAREs. 
%
%For clarity, we consider the first part of the bit vector as Q-SNAREs and rest as R-SNAREs.
%
%An edge in the system with $M$ = 4 molecules is interpreted as a label [$M_{1}, M_{2}, M_{3}, M_{4}$]. 
%
%Where $M_{1}, M_{2}, M_{3}$ are Q-SNAREs while $M_{4}$ is a R-SNARE.
%
%Also, each $M_i$ can be either 0 or 1, representing the presence or absence of the $i^{th}$ molecule.
%
%For example, the variable \texttt{bvEdge}[1][2][1] = 1001, represents the first edge between node 1 and node 2 with molecule $M_{1}$ (Q-SNARE) and $M_{4}$ (R-SNARE) present. 
%
%Similarly, a node in the graph is also labelled as a bit vector of length $M$.   
%
%We leave the graph completely arbitrary and add the constraint to enforce that the graph is k-connected. 
%
%
The value of the bit vector variable describe the presence of the corresponding molecules at that particular location. 
%
For clarity, we consider the first part of the bit vector as Q-SNAREs and rest as R-SNAREs.
%
For an edge in the system with $M$ = 4 molecules, the variable \texttt{bvEdge}[0][2][1] = $1001$, means the molecule  $M_{1}$ (Q-SNARE) and $M_{4}$ (R-SNARE) are present on the first edge between node 2 and node 1.

To specify the stability condition, we use nondeterminism to perform enumeration.
%
%Nondeterminism can be viewed as an oracle that guesses a correct path for a specific $m$ if it exists. 
%
%We then use this path as a witness to ensures that stability is maintained for each outgoing molecule $m$ on every edge $E$.
%
%
Note that the enumeration is done only for an unwinding depth $w$.
%
The encoding presented below is a simpler version of the implementation. 
\subsubsection{Constraints for stability condition}
%
A path between node $i$ and $j$ in the graph is a $m$-$path_{ij}$, if molecule $m$ is present on edges along that path.
%
%We use {\npath} to represent the required path to encode the stability condition in VTS.
%
%A naive implementation is shown in~\ref{eq:reachbmc1} where 
We use {\npath}$(m, i, j, w)$ to represent the presence of a $m$-$path_{ij}$ of maximum $w$ length.
%there is a edge between $i^{\prime \prime}$ and01
%$(i \neq i^{\prime \prime})$ with $m$ present, and j is $m$-reachable from $i'$ in less than $p-1$ steps.
%
% \begin{align}
%   NOracle (m, i, j, w) = \exists \, path[h1, ..., hn]: \\ 2 \leq |path| \leq  w + 2  \, \land h1 = i \, \land  hn = j \, \land \\ 
%   \bigvee\limits_{k = 1}^{|path|- 1} \bigvee\limits_{q}  E[hk] [hk+1][q][m] 
%   \tag{S1}\label{eq:reach1}
% \end{align}
%The NOracle can be thought of as a disjunction of $m$-path of increment length. 
%
$\npath(m, i, j, w)$ performs enumeration of every $m$-$path_{ij}$ of at most $w$ length. 
%
Starting with exploring $m$-${paths}_{ij}$ of length 1; a direct edge between $i$ and $j$ with molecule $m$ present, until every $m$-${paths}_{ij}$ of length $w-1$ is explored.
%by performing series of node hops till final node destination j is reached.
%Starting with $m$-${paths}_i^j$ of length 1 which is a direct edge between $i$ and $j$ with molecule $m$ present on the edge or performing series of node hops till final node destination j is reached.

@@

\subsubsection{The key difference in the reachability encoding}
The key improvement of the $\fsmt$ search encoding over the $\fbmc$ is the
encoding of reachability.
% which was done using an enumeration of paths.
%
In the $\fbmc$ encoding, reachability is encoded by enumerating all the possible paths using non-determinism.
%
This encoding is non-optimal in the sense that in the worst case an exponential number of paths needs to be explored.
%
Whereas in the $\fsmt$ encoding, reachability is encoded in two different
ways in constraints $\texttt{Paths}(r)$ and $\texttt{FReach}(d,r')$.
%
The reachability is recursively defined in~$\texttt{FReach}(d,r')$ and has
trivial solutions by making all $r'$s true.
%
However, the trivial solutions are disallowed by constraint~$\texttt{Connected}(r')$ and we find
only the solutions that captures the evidence of non-reachability.
%
On the other hand, we have added length of paths in our reachability encoding in constraint~$\texttt{Paths}(r)$, which needs relatively more auxiliary variables.
%
This is because constraint~$\texttt{Loop}(r)$ has only positive
occurrences of the reachability variables and if we had defined
$r$s in~$\texttt{Paths}(r)$ without paths,
the circular dependencies in the recursive definitions of $r$s
may have resulted in spurious satisfying assignments that
do not encode reachability.
%
By adding the path length, we break the circular dependencies, the
constraint remains polynomial in size, and satisfying assignments only
correspond to correct reachability.

@@

We have encoded the following variants of VTSs.
%
The variants are due to different combinations of constraints on the
activity of molecules on the nodes and edges.
%
\begin{enumerate}
\item Every present molecule is considered to be active.
\item Activity of molecules on the nodes is based on Boolean function of presence of other molecules. 
\item Activity of molecules on the edges is based on Boolean function of presence of other molecules
\item Activity of molecules both on the edges and nodes is based on Boolean function of presence of other molecules.
\item Activity of molecules on the edges is driven by pairing inhibition.
\item Activity of molecules on the nodes is based on Boolean function of presence of other molecules and on edge by pairing inhibition.
\end{enumerate}
%
% For the version A, every present molecule is in active state.
%

>>>>

%\externaldocument{sat-table}
%\externaldocument{smt-table}
%\externaldocument{qbf-table}
\noindent In the previous section, we have discussed in total four problems, $\fbmc$ search, $\fsmt$ search, $\fqbf$ search and synthesis problem.
%
In this section, we will present encodings for each of them.
%
Although the underlying technique for $\fbmc$ and $\fsmt$ search varies, the encoding is very similar. 
%
So, instead of presenting serially the complete encoding of both, we will start with the complete $\fsmt$ search encoding and present only the relevant part of the $\fbmc$ encoding to show a contrast.
%contrast with the $\fbmc$ encoding.

@@

%\subsection{Boolean satisfiability of the search problem}
\noindent We translate the search problem into a Boolean satisfiability
problem and use SMT solvers to find the satisfying VTSs.
%
We will first present the variables used to encode the
VTSs and the properties.
%
Afterward, we will present the constraints corresponding to the
properties.

\subsubsection{Variables for VTS description}
%
We assume that the size of the graph is $\nu$ and number of molecules is
$\mu$.
%
Furthermore, we also limit the maximum number of $\pi$ of edges present
between two nodes.
%
Here, we list the vector of Boolean variables and uninterpreted function symbols that encode the VTSs.
\begin{enumerate}

@@

\subsubsection{Constraints on presence, activity of the molecule, and pairing matrix}
%
We need the following constraints \texttt{EdgeC},
\texttt{ActivityC} and \texttt{PairingC} to encode the basic structure of VTSs.
%
For an edge to exist it should have one molecule present. 
%
\begin{align*}
  \texttt{PresentE} = \bigwedge\limits_{i,j,q} (\bigvee_m e_{i,j,q,m} )\limplies e_{i,j,q}
\end{align*}
The edge labels are subset of the node labels of the source and target nodes.
\begin{align*}
  \texttt{LabelE} = \bigwedge\limits_{i,j,q,m} e_{i,j,q,m} \limplies (n_{i,m} \land n_{j,m} )
\end{align*}
Self edges are not allowed. 
\begin{align*}
  \texttt{SelfE} = \bigwedge\limits_{i,q} \neg e_{i,i,q}
\end{align*}
\noindent We define \texttt{EdgeC} to collect all the edge related constraints. 
%
\texttt{EdgeC} states that each edge has at least one molecule,
 edge labels are consistent with node labels and there are no self-loops.    
\begin{align*}
  \texttt{EdgeC} = \texttt{PresentE} \land \texttt{LabelE}  \land \texttt{SelfE} 
\end{align*}
\noindent If a molecule is active on an edge, it should be present on the edge.
%
\begin{align*}
  \texttt{ActiveMolE} = \bigwedge\limits_{i,j,q,m} b_{i,j,q,m} \limplies e_{i,j,q,m}\label{eq:f1}
\end{align*}
A molecule should be present to be active on a node.  
\begin{align*}
    \texttt{ActiveMolN} = \bigwedge\limits_{i,m} a_{i,m} \limplies n_{i,m}
\end{align*}
\noindent We define \texttt{ActivityC} to collect all the molecule activity related constraints. 
\texttt{ActivityC} states that all active molecules are present.
\begin{align*}
  \texttt{ActivityC} = \texttt{ActiveMolE} \land \texttt{ActiveMolN} 
\end{align*}
\noindent We divide the pairing matrix into Q-SNAREs and R-SNAREs blocks and set the dependency bits of R-SNAREs to R-SNAREs in the pairing matrix to be all 0's.
%It reduces the search space.
% Condition on $p_{kk'}$. \ashu{why???} .
%\begin{align}
%  \bigwedge\limits_{(x < M/2 \, \land  \, y < M/2) \lor  (x >= M/2 \land y >= M/2)} \neg p(x,y)
%  \tag{V6}\label{eq:c3}
%\end{align}

@@

\subsubsection{Well-fused constraints}
Constraint \texttt{Fusion1} encodes that each edge must fuse with
its destination node.
%
Constraint \texttt{Fusion2} encodes that each edge should not
be able to potentially fuse with any node other than its destination node.
\begin{align*}
\texttt{Fusion1} =\;&
\bigwedge\limits_{i,j,q} e_{i,j,q} \limplies
\bigvee_{{m_1,m_2,m_3,m_4}} (\Land_{l=1}^4 ( b_{i,j,q,m_l} \lor a_{j,m_l} ) \land 
\Lor\limits_{l=1}^4 b_{i,j,q,m_l} \land\\
& \qquad \qquad \qquad \qquad \Lor\limits_{l=1}^4 a_{j,m_l}\land p_{\{m_1,m_2,m_3,m_4\}})
\\
\texttt{Fusion2} =\;&
\bigwedge\limits_{\mathclap{i,j,q,m_1,m_2,m_3, l \in \{1,..,3\}}} b_{i,j,q,m_1} \land ..\land b_{i,j,q,m_l} \limplies \\
& \hspace{2cm} \neg 
\bigvee_{\mathclap{j \neq j^{\prime}, m_{l+1}^{\prime},..,m_{4}^{\prime}}} ( a_{j^{\prime},m_{l+1}^{\prime}} \land .. \land a_{j^{\prime},m_4^{\prime}} \land p_{\{m_1,..,m_l,m^{\prime}_{l+1},..,m^{\prime}_4\}})
\\
\end{align*}
We define \texttt{FusionC} \change{Use of the word collect looks weird! change it.}to collect all the fusion related constraints.
% 
\texttt{FusionC} states that every edge in the graph is well-fused.
\begin{align*}
\texttt{FusionC} = \texttt{Fusion1} \land \texttt{Fusion2} 
\end{align*}
% Fusion rules consist of two different mechanisms.
% \begin{enumerate}
% \item  A \textbf{pairing mechanism} which determines compatible Q-R pairs on vesicles and compartments that can cause fusion.
% \item \textbf{regulatory mechanisms} on the edges and nodes (possibly distinct) which regulates molecules activity based on the presence/absence of other molecules on the corresponding node or edge.
% \end{enumerate}
% Boolean constraint~\eqref{eq:fuse1} and~\eqref{eq:fuse2} are used to encode well-fused
% property.

@@

\subsubsection{Regulation on nodes and edges (Activity function)}
We are considering several models of VTS that differ in types of regulations on the molecule both on the node and on the edge.  
%constraints on the activity of molecules.
%
%We will present~\eqref{eq:ann}-~\eqref{eq:aep} that encodes
%the varying constraints.
%
The activity of the molecules is controlled by activity functions.
%
We will have encoded five variations of the Boolean functions based on the biological discussion.
%
The concrete functions encoding is presented below.
%
We will use \texttt{NodeFun}$_m$ and \texttt{EdgeFun}$_m$ for node and edge regulations respectively.
%

The following constraint encodes no conditions on activities on nodes,
i.e., all the present molecules on the nodes are active.
\begin{align*}
\texttt{FunNoN} =& \bigwedge\limits_{i,m} n_{i,m} = a_{i,m}  
\end{align*}
The following constraint encodes that activity of a molecule $m$ on the node is
defined by a Boolean function $\nodef_m$ of presence of molecules present on that node.
\begin{align*}
\texttt{FunBfN} =& \bigwedge\limits_{i,m} n_{i,m} \limplies a_{i,m} =  \nodef_m (n_{i,1},\dots,n_{i,\mu}) 
\end{align*}
The following constraint encodes no conditions on activities on edges,
i.e., all the present molecules on the edges are active.
\begin{align*}
 \texttt{FunNoE} =& \bigwedge\limits_{i,j,q,m} e_{i,j,q,m} = b_{i,j,q,m}
\end{align*}
The following constraint encodes that activity of a molecule $m$ on the edge is defined by a Boolean function $\edgef_m$ of the presence of molecules present on that edge.
\begin{align*}
 \texttt{FunBfE} =& \bigwedge\limits_{i,j,q,m} e_{i,j,q,m} \limplies b_{i,j,q,m} = \edgef_k(e_{i,j,q,1}, .., e_{i,j,q,\mu} )
\end{align*}
%
The following constraint encodes that the activity of the molecules on
edges is defined by inhibition by other molecules based on the pairing
matrix. 
\begin{align*}
\texttt{FunPairE} =&	\begin{aligned}[t]
\bigwedge\limits_{i,j,q, m}  [e_{i,j,q,m} \land \bigwedge_{m_l \neq m,l \in \{1,..,3\}} e_{i,j,q,m'} \land p_{m, m_l}] \limplies\\
	(\neg b_{i,j,q,m} \land \bigwedge_{\mathclap{m_l \neq m,l, p_{m, m_l} \in \{1,..,3\}}} \neg b_{i,j,q,m_l})\end{aligned}\\
\end{align*}
%
We denote the particular choice of node and edge function combined by \texttt{ActivityFun}.
\begin{align*}
\texttt{ActivityFun} = \texttt{NodeFun}_m \land \texttt{EdgeFun}_m
\end{align*}
\subsubsection{Constraints for stability condition}
%
We use $m$-reachability to encode the stability condition in VTSs.
%
The following constraint recursively encodes that node $j$ is $m$-reachable from node $i$ in less than $p$ steps
if either there is a direct edge between $i$ and $j$ with $m$ present on the edge or there is a edge between $i^{\prime \prime}$ and
$(i \neq i^{\prime \prime})$ with $m$ present, and j is $m$-reachable from $i'$ in less than $p-1$ steps.
%
\begin{align*}
  \texttt{Paths}(r) &= \bigwedge\limits_{i,j,m,p} r_{i,j,m,p} \limplies (\bigvee_{q} \, e_{i,j,q,m} \lor \bigvee_{i\neq i^{\prime}} ( \, \bigvee_{q} e_{i,i^{\prime},q,m}) \land r_{i^{\prime},j,m,p-1} )
\end{align*}
Now we can encode stability using the reachability variables.
and say if there is an $m$-edge between $i$ and $j$, there is
$m$-reachable path from $j$ to $i$.
\begin{align*}
 \texttt{Loop}(r) &= \bigwedge\limits_{i,j,m} (\bigvee_{q} e_{i,j,q,m}) \limplies r_{j,i,m,\nu}
\end{align*}
We define \texttt{StabilityC} to collect all the stability related constraints.
% 
\texttt{StabilityC} states that every molecule is moving in a cycle.
\begin{align*}
\texttt{StabilityC} = \texttt{Paths}(r) \land \texttt{Loop}(r) 
\end{align*}
% Boolean variable
% F3 and F2 are used to model these constraints.

@@

\noindent Our first attempt was to encode the existential condition as a $\fbmc$ search \unsure{Should we cite it here or not?}problem~\cite{shukla2017discovering}.
%
%In this section we will present the $\fbmc$ encoding of the same.
%The basic idea is to search the solution only upto a certain depth $w$. 
%
%\begin{align*}
%\npath(m, i, j, w) = \bigvee\limits_{l = 1}^{w-1} (\texttt{Length} (m-path_{ij}) == l)
%\end{align*}
%
%
%\begin{align*}
%\bigwedge\limits_{i,j,m} (\bigvee_{q} (\texttt{bvEdge}[i][j][q] \unrhd m) == 1) \limplies \npath(m, j, i, w)
%\end{align*}
%Although the techniques used for $\fbmc$ and $\fsmt$ is very different because we are 
%In $\fbmc$ we only partially unroll the program to check for the solutions within that subset. 
%in the case of $\fbmc$, the encoding is not considerably different.
% 
The main differences between $\fsmt$ and $\fbmc$ are the representation of variables and the encoding of 
%I n comparison to $\fsmt$, the $\fbmc$ search encodes 
reachability (both for the stability and connectivity).  
%]
Here, we will only present the stability encoding.
%the rest of the encodings are similar to the one presented in the previous section, Section~\ref{enc:smt}.

@@

% with each bit corresponding to the presence or absence of the molecule. 
%
%In the $\fbmc$ encoding, variables are represented as an array of bit vectors of size $M$. 
%
%We initialize the system with a fixed number of node size (N) and distinct molecular types (M). 
%%
%We use a two-dimensional array variable \textit{Graph}[N][N] to represent the transport graph. The graph variable is set to x, Graph[i][j] = x, if there are exactly x edges between node i and j. 
%\begin{align*}
%graph [i] [j] = x.
%\end{align*}
%
%For example, every edge is represented as a bit vector of size $M$ with each bit corresponding to the presence or absence of the molecule. 
%
%Bit vector \texttt{bvEdge}$[i][j][q]$ represent the $q$th edge between node $i$ and $j$.
%
%The index of the bit vector implicitly determines the corresponding molecule we are referring to in the edge.
% 
%The molecule set is divided into Q and R-SNAREs. 
%
%For clarity, we consider the first part of the bit vector as Q-SNAREs and rest as R-SNAREs.
%
%An edge in the system with $M$ = 4 molecules is interpreted as a label [$M_{1}, M_{2}, M_{3}, M_{4}$]. 
%
%Where $M_{1}, M_{2}, M_{3}$ are Q-SNAREs while $M_{4}$ is a R-SNARE.
%
%Also, each $M_i$ can be either 0 or 1, representing the presence or absence of the $i^{th}$ molecule.
%
%For example, the variable \texttt{bvEdge}[1][2][1] = 1001, represents the first edge between node 1 and node 2 with molecule $M_{1}$ (Q-SNARE) and $M_{4}$ (R-SNARE) present. 
%
%Similarly, a node in the graph is also labelled as a bit vector of length $M$.   
%
%We leave the graph completely arbitrary and add the constraint to enforce that the graph is k-connected. 
%
%
The value of the bit vector variable describes the presence of the corresponding molecules at that particular location. 
%
For clarity, we consider the first part of the bit vector as Q-SNAREs and rest as R-SNAREs.
%
For an edge in the system with $M$ = 4 molecules, the variable \texttt{bvEdge}[2][1][0] = $1001$, means the molecule  $M_{1}$ (Q-SNARE) and $M_{4}$ (R-SNARE) are present on the first edge between node 2 and node 1.

To specify the stability condition, we use \improve{Nondet usage is not clear in the encoding. improve.}nondeterminism to perform enumeration.
%
%Nondeterminism can be viewed as an oracle that guesses a correct path for a specific $m$ if it exists. 
%
%We then use this path as a witness to ensures that stability is maintained for each outgoing molecule $m$ on every edge $E$.
%
%
Note that the enumeration is done only for an unwinding depth of $w$.
%
The encoding presented below is a simpler version of the implementation. 
\subsubsection{Constraints for stability condition}
%
A path between node $i$ and $j$ in the graph is an $m$-$path_{ij}$ if molecule $m$ is present on edges along that path.
%
%We use {\npath} to represent the required path to encode the stability condition in VTS.
%
%A naive implementation is shown in~\ref{eq:reachbmc1} where 
We use {\npath}$(m, i, j, w)$ to represent the presence of a $m$-$path_{ij}$ of maximum $w$ length.
%there is a edge between $i^{\prime \prime}$ and01
%$(i \neq i^{\prime \prime})$ with $m$ present, and j is $m$-reachable from $i'$ in less than $p-1$ steps.
%
% \begin{align}
%   NOracle (m, i, j, w) = \exists \, path[h1, ..., hn]: \\ 2 \leq |path| \leq  w + 2  \, \land h1 = i \, \land  hn = j \, \land \\ 
%   \bigvee\limits_{k = 1}^{|path|- 1} \bigvee\limits_{q}  E[hk] [hk+1][q][m] 
%   \tag{S1}\label{eq:reach1}
% \end{align}
%The NOracle can be thought of as a disjunction of $m$-path of increment length. 
%
$\npath(m, i, j, w)$ performs enumeration of every $m$-$path_{ij}$ of at most $w$ length. 
%
Starting with exploring $m$-${paths}_{ij}$ of length 1; a direct edge between $i$ and $j$ with molecule $m$ present, until every $m$-${paths}_{ij}$ of length $w-1$ is explored.\change{m-path looks - sign in the align}
%by performing series of node hops till final node destination j is reached.
%Starting with $m$-${paths}_i^j$ of length 1 which is a direct edge between $i$ and $j$ with molecule $m$ present on the edge or performing series of node hops till final node destination j is reached.

@@

\subsubsection{The key difference in the reachability encoding}
The key improvement of the $\fsmt$ search encoding over the $\fbmc$ is the
encoding of reachability.
% which was done using an enumeration of paths.
%
In the $\fbmc$ encoding, reachability is encoded by enumerating all the possible paths using \improve{Do not show up in the encoding}nondeterminism.
%
This encoding is non-optimal in the sense that in the worst case an exponential number of paths needs to be explored.
%
Whereas in the $\fsmt$ encoding, reachability is encoded in two different
ways in constraints $\texttt{Paths}(r)$ and $\texttt{FReach}(d,r')$.
%
The reachability is recursively defined in~$\texttt{FReach}(d,r')$ and has
trivial solutions by making all $r'$s true.
%
However, the trivial solutions are disallowed by constraint~$\texttt{Connected}(r')$ and we find
only the solutions that capture the evidence of non-reachability.
%
On the other hand, we have added the length of paths in our reachability encoding in constraint~$\texttt{Paths}(r)$, which needs relatively more auxiliary variables.
%
This is because constraint~$\texttt{Loop}(r)$ has only positive
occurrences of the reachability variables and if we had defined
$r$s in~$\texttt{Paths}(r)$ without paths,
the circular dependencies in the recursive definitions of $r$s
may have resulted in spurious satisfying assignments that
do not encode reachability.
%
By adding the path length, we break the circular dependencies, the
constraint remains polynomial in size, and satisfying assignments only
correspond to correct reachability.

@@

We have encoded the following variants of VTSs.
%
The variants are due to different combinations of constraints on the
activity of molecules on the nodes and edges.
%
\begin{enumerate}
\item Every present molecule is considered to be active.
\item Activity of molecules on the nodes is based on Boolean function of the presence of other molecules. 
\item Activity of molecules on the edges is based on Boolean function of the presence of other molecules
\item Activity of molecules both on the edges and nodes is based on Boolean function of the presence of other molecules.
\item Activity of molecules on the edges is driven by pairing inhibition.
\item Activity of molecules on the nodes is based on Boolean function of the presence of other molecules and on edge by pairing inhibition.
\end{enumerate}
%
% For the version A, every present molecule is in active state.
%

In the Table~\ref{tab:var-grph}, we present the constraints involved in each version.
%
The column two of the table shows that the constraints $\texttt{EdgeC}$, $\texttt{ActivityC}$, $\texttt{PairingC}$, $\texttt{FusionC}$, $\texttt{StabilityC}$ and $\texttt{Connected}(k)$ are present for every variant.
%
The last two columns of the table lists the activity constraints that are different among the variants.
%
%
%One of the restriction is where the activity of the present molecule is dependent on the presence of the other molecules.
%
%For example in version B, D, F activity of the molecules on the node is a boolean function of the presence of other molecules on that node; Anb.
%
%Similarly for the case of the edge in version C and D; Aeb.
%
%In the case of versions F the activity of the molecules on the edges
%is described by pairing matrix; Aep.
